Convex Surrogates via Polyhedral Losses

### 2019-01-31 cvx-fin.tex
tiny bit of writing at end of 4.0. bo@nyc
mode lower bound. bo@nyc

\begin{corollary}\label{cor:d-embed-implies-d-construction}
  If a property is $d$-embeddable, then each of its level sets has a $d$-construction.
  \bo{Is it possible that some uniqueness of the construction is also true?}\jessie{Up to rotation, scaling, and translation, possible but I wouldn't think so.  ex: Abstain(1/2) on 3 outsomes can be embedded in $\reals^2$ with both an equilateral and isosceles triangle}
\end{corollary}
\begin{proof}
  Suppose the level set is represented as $\Gamma_r = \{p : Bp \geq 0\}$ with $B \in \reals^{k \times n}$ for the minimum $k$ (i.e. the number of facets of the polytope).
  Let $u \in \reals^d$ be its embedding point.
  The convex expected loss is minimized at $u$ if and only if it has the zero vector as a subgradient at $u$, i.e.
    \[ p \in \Gamma_r \qquad \iff \qquad 0 \in \partial \langle L(u), p \rangle . \]
  Subgradient sets satisfy the following two properties: $\partial(f_1 + f_2) = \partial f_1 + \partial f_2$ where the right side is the Minkowski sum of the subgradient sets (i.e. the set of sums of all pairs); and $\partial(\alpha f_1) = \alpha \partial f_1$.
  So
    \[ \partial \langle p , L(u) \rangle = \sum_{y \in \Y} p_y \partial L(u)_y \]
  where the right side is a Minkowski sum of subgradient sets.
  We now consider the halfspaces defining this set.
  Define the $y$th support function as $H_y(v) = \max_{w \in \partial L(u)_y} \langle w, v \rangle$ to obtain that $w \in \partial \langle p, L(u) \rangle$ if and only if, for all unit vectors $v$,
   \[ \langle v, w \rangle \leq \sum_{y \in \Y} p_y H_y(v) . \]
  In particular, the embedding point is optimal for $p$ if and only if this is satisfied for $w=0$.
  So, letting $B_i$ be the $i$th column of $B$, we have
    \[ \Gamma_r = \{p : \sum_y p_y B_y \geq 0\} = \{p : \sum_y p_y H_y(v) \geq 0 ~ (\forall v, \|v\|=1)\} . \]
  \bo{This is the part I'm still not perfectly happy with}
  These sets are equal and full-dimensional (as the property is finite and non-redundant and $B$ is minimal).
  So there exist $k$ unit vectors $v_1,\dots,v_k \in \reals^d$ such that $H_y(v_i) = B_{iy}$.
  Collected as rows of $V$, we obtain that each $(V,B_y)$ is a tight halfspace representation, as it contains $\partial L(u)_y$ and $H_y(v_i) = B_{iy}$, meaning there exists $w \in \partial L(u)_y$ such that $\langle w, v_i \rangle = B_{iy}$.
\end{proof}
This provides the following recipe for both constructing losses and proving lower bounds.
First, take any level set and represent it as $\{p : Bp \geq 0\}$.
Now if $B$ has $k$ columns, search for a collection of $k$ unit vectors in $\reals^d$.
For any candidate $V$ (whose rows are the unit vectors), a computer program can automatically verify whether the representations $\{(V,B_y)\}_{y \in \Y}$ are tight.
In particular, if any $(V,B_y)$ is infeasible (i.e. there does not exist any $w$ satisfying all constraints $\langle v_i, w \rangle \leq B_{iy}$), then $V$ is not a $d$-construction.
If $(V,B_y)$ is feasible but some constraints are redundant, i.e. it is not tight, then again we do not have a $d$-construction.

A lower bound can be proven by showing, for any level set, that no collection of $k$ unit vectors can form a $d$-construction.

@@

@@

\begin{theorem}
  The mode with $|\Y|=n$ is not $(n-2)$-embeddable.
\end{theorem}

>>>>

\begin{corollary}\label{cor:d-embed-implies-d-construction}
  If a property is $d$-embeddable, then each of its level sets has a $d$-construction.
  \bo{Is it possible that some uniqueness of the construction is also true?}\jessie{Up to rotation, scaling, and translation, possible but I wouldn't think so.  ex: Abstain(1/2) on 3 outsomes can be embedded in $\reals^2$ with both an equilateral and isosceles triangle}
\end{corollary}
\begin{proof}
  Suppose the level set is represented as $\Gamma_r = \{p : Bp \geq 0\}$ with $B \in \reals^{k \times n}$ for the minimum $k$ (i.e. the number of facets of the polytope).
  Let $u \in \reals^d$ be its embedding point.
  The convex expected loss is minimized at $u$ if and only if it has the zero vector as a subgradient at $u$, i.e.
    \[ p \in \Gamma_r \qquad \iff \qquad 0 \in \partial \langle L(u), p \rangle . \]
  Subgradient sets satisfy the following two properties: $\partial(f_1 + f_2) = \partial f_1 + \partial f_2$ where the right side is the Minkowski sum of the subgradient sets (i.e. the set of sums of all pairs); and $\partial(\alpha f_1) = \alpha \partial f_1$.
  So
    \[ \partial \langle p , L(u) \rangle = \sum_{y \in \Y} p_y \partial L(u)_y \]
  where the right side is a Minkowski sum of subgradient sets.
  We now consider the halfspaces defining this set.
  Define the $y$th support function as $H_y(v) = \max_{w \in \partial L(u)_y} \langle w, v \rangle$ to obtain that $w \in \partial \langle p, L(u) \rangle$ if and only if, for all unit vectors $v$,
   \[ \langle v, w \rangle \leq \sum_{y \in \Y} p_y H_y(v) . \]
  In particular, the embedding point is optimal for $p$ if and only if this is satisfied for $w=0$.
  So, letting $B_i$ be the $i$th column of $B$, we have
    \[ \Gamma_r = \{p : \sum_y p_y B_y \geq 0\} = \{p : \sum_y p_y H_y(v) \geq 0 ~ (\forall v, \|v\|=1)\} . \]
  \bo{This is the part I'm still not perfectly happy with}
  These sets are equal and full-dimensional (as the property is finite and non-redundant and $B$ is minimal).
  So there exist $k$ unit vectors $v_1,\dots,v_k \in \reals^d$ such that $H_y(v_i) = B_{iy}$.
  Collected as rows of $V$, we obtain that each $(V,B_y)$ is a tight halfspace representation, as it contains $\partial L(u)_y$ and $H_y(v_i) = B_{iy}$, meaning there exists $w \in \partial L(u)_y$ such that $\langle w, v_i \rangle = B_{iy}$.
\end{proof}
This provides the following recipe for both constructing losses and proving lower bounds.
First, take any level set and represent it as $\{p : Bp \geq 0\}$.
Now if $B$ has $k$ rows, search for a collection of $k$ unit vectors in $\reals^d$.
For any candidate $V$ (whose rows are the unit vectors), a computer program can automatically verify whether the representations $\{(V,B_y)\}_{y \in \Y}$ are tight.
In particular, if any $(V,B_y)$ is infeasible (i.e. there does not exist any $w$ satisfying all constraints $\langle v_i, w \rangle \leq B_{iy}$), then $V$ is not a $d$-construction.
If $(V,B_y)$ is feasible but some constraints are redundant, i.e. it is not tight, then again we do not have a $d$-construction.

@@

A lower bound can be proven by showing, for any level set, that no collection of $k$ unit vectors can form a $d$-construction.
We will give examples of this in $2$ dimensions next.

@@

\begin{theorem}[\bo{cite}]
  The mode with $|\Y|=n$ is not $(n-2)$-embeddable.
\end{theorem}
\begin{proof}
  Call the outcomes $\Y = \{1,\dots,n\}$.
  We will pick the level set for the report $r=1$ and show it has no $d$-representation.
  The level set can be represented as $\{p : p_1 \geq p_i ~ (\forall i \neq 1)\}$, or the set of $p$ such that $Bp \geq 0$ with
  \[ B = \begin{pmatrix}
       1  & -1 &  0 & \cdots & 0  \\
       1  &  0 & -1 & \cdots & 0  \\
          &    &    & \vdots &    \\
       1  &  0 &  0 & \cdots & -1
     \end{pmatrix} .
  \]
  Here $B$ has $n-1$ rows, one for each $i \neq 1$.

  A $d$-representation is therefore a $V$ with $n-1$ rows, each a unit vector $v_i$ in $\reals^d$.
  Suppose $d \leq n-2$, i.e. there are at least $d+1$ rows.
  Now consider the halfspace representation given by the second column of $B$, namely $T = \{ w : Vw \leq (-1, 0, \dots, 0)^{\intercal}\}$.
  The first $d$ rows of $V$ form a (not-necessarily-orthonormal, but normal) basis for $\reals^d$.
  Therefore, we can write the $d+1$st row as $v_{d+1} = \alpha_1 v_1 + \cdots + \alpha_d v_d$.
  So if $w \in T$, then $\langle v_1,w \rangle \leq -1$ and $\langle v_i,w \rangle \leq 0$ for all $i=2,\dots,d$.
  Taking the convex combination, we obtain $\langle v_{d+1}, w \rangle \leq -\alpha_1$.
  If $\alpha_1$ is positive, this implies there is slack in the $d$th constraint, so the halfspace representation cannot be tight.
  If $\alpha_1$ is positive, then we pick some other pair of vectors $i \neq j$ rather than $1,d+1$, such that $v_j = \sum_{i'\neq j} \alpha_{i'} v_{i'}$ with $\alpha_{i'} > 0$.
  \bo{I claim there must exist two such vectors in this collection ... need to prove it though!}
  Choosing the $i+1$st column of $B$ as a constraint gives the contradiction as above.
\end{proof}
To recall what is happening inside this proof: we chose the first level set, which gave us a matrix $B$.
The first column was all-$1$, which corresponds to a constraint on the subgradient set of the loss on outcome $1$ for report $1$, which is not an interesting constraint.
The other constraints gave a contradiction if $d$ is too small.
\bo{More to say?}
